### 2011-09-21 expectationPeopleTable.tex
first table version

\begingroup
\newcommand{\add}{\hspace{0pt}}

\definecolor{white}{rgb}{1,1,1}
\definecolor{mygray}{rgb}{0.7,0.7,0.7}
\definecolor{yourgray}{rgb}{0.4,0.4,0.4}
\definecolor{black}{rgb}{0.0,0.0,0.0}

\newdimen\qdx
\newdimen\qda
\newdimen\qdb
\newdimen\qd
\def\rrrr#1#2#3#4#5#6{\qd=#4 % length of bar for 1.0
\qdx=\qd\multiply\qdx by 5\divide\qdx by 4
\qda=\qd
\qdb=\qd
\multiply\qda by #1\divide\qda by #3\multiply\qdb by #2\divide\qdb by #3\advance\qdb by -\qda
    \leavevmode\hbox to \qdx{\hfil\vbox{%
    \hbox{\vrule\vbox{\hrule\hbox to 1\qd
            {\vrule depth0pt height#6 width \qda#5\vrule depth0pt height#6 width \qdb\hfill}\hrule}\vrule}
    }\hfil}}
\def\rrr#1#2#3#4{\rrrr{#1}{#2}{#3}{0.19cm}{#4}{1.5ex}}

\def\w{\rrr{0}{1}{1}{\color{white}}}
\def\l{\rrr{0}{1}{1}{\color{mygray}}}
\def\g{\rrr{0}{1}{1}{\color{yourgray}}}
\def\b{\rrr{0}{1}{1}{\color{black}}}

\begin{table*}[tb]
\centering
%\footnotesize
%\scriptsize
\begin{tabular}{l@{\hspace{-3cm}}r@{\hspace{2pt}}c}
\toprule
& Expectations &People\\
\midrule
%&&& $p_1$ & $p_2$& $p_3$& $p_4$& $p_5$& $p_6$& $p_7$& $p_8$& $p_9$& $p_{10}$& $p_{11}$& $p_{12}$& $p_{13}$& $p_{14}$& $p_{15}$& $p_{16}$& $p_{17}$& $p_{18}$& $p_{19}$& $p_{20}$& $p_{21}$& $p_{22}$& $p_{23}$& $p_{24}$& $p_{25}$& $p_{26}$& $p_{27}$& $p_{28}$& $p_{29}$& $p_{30}$& $p_{31}$& $p_{32}$& $p_{33}$& $p_{34}$& $p_{35}$& $p_{36}$\\
%&&&p&p&p&p&p&p&p&p&p&p&p&p&p&p&p&p&p&p&p&p&p&p&p&p&p&p&p&p&p&p&p&p&p&p&p&p\\
%\midrule
\textbf{Technical - Core} 
&Write bug-fix code&  \w\w\l\w\g\b\b\b\g\w\w\w\w\w\l\l\l\l\l\l\l\l\l\l\l\l\l\l\l\l\l\l\l\l\l \\
&Maintain legacy code&  \w\w\l\w\g\b\b\b\g\w\w\w\w\w\l\l\l\l\l\l\l\l\l\l\l\l\l\l\l\l\l\l\l\l\l \\
&Write new-feature code&\\
&Write customization (or script) code&\\
%& \\
%
%
\midrule
\textbf{Technical - Periphery}
& Write build code&\\
& Design user interface&\\
& Write user documentation&\\
& Bundle and prepare a release&\\
& Write technical documentation&\\
&Write low-level (or unit) test code&\\
& Write experimental (or exploratory) code&\\
& Write high-level (or integration) test code&\\
& Write configuration and implementation code&\\
& Write and maintain tools to aid in development&\\
%& \\
%
%
\midrule
\textbf{Technical - High level}
& Elicit requirements&\\
& Document requirements&\\
& Design system architecture&\\
& Determine if product meets quality goals&\\
& Set roadmap for system (prioritize requirements)&\\
& Provide domain context to non-domain experts (user proxy)&\\
%& \\
%
%
\midrule
\textbf{Technical - Expertise}
& Tackle the toughest technical issues and bugs&\\
& Provide historical perspective on design decisions&\\
& Provide guidance (mentorship) to newcomers on technical issues&\\
%& \\
%
%
\midrule
\textbf{Technical - Management}
& Monitor low-level task (ticket) progress&\\
& Monitor high-level task (project) progress&\\
& Determine lifecycle process and practices&\\
& Document (notify) on one's own progress&\\
& Assign low-level tasks (tickets) to individuals&\\
& Assign high-level tasks (projects) to individuals&\\
& Coordinate a product release or implementation&\\
& Provide estimates on remaining or potential tasks&\\
& Connect an issue with the person with appropriate expertise&\\
& Handle obstacles and interruptions for technical teams (PoC)&\\
%& \\
%
%
\midrule
\textbf{Organizational - External Interface}
& Obtain new users&\\
& Service current users&\\
& Receive and respond to user feedback&\\
& Obtain and maintain sources of funding&\\
& Provide domain legitimacy to organization&\\
& Provide technical legitimacy to organization&\\
& Advocate for the resolution of user/customer problems&\\
& Represent the organization's interests to funders/customers&\\
%& \\
%
%
\midrule
\textbf{Organizational - Management}
& Deal with clerical issues&\\
& Deal with financial issues&\\
& Deal with personnel issues&\\
& Deal with otherwise unattended issues&\\
& Deal with system administration issues&\\
& Set up and modify organizational structure&\\
& Communicate organizational status and news&\\
%& \\
%
%
\midrule
\textbf{Organizational - Strategic}
& Determine long-term vision for products&\\
& Determine emphasis on new market areas&\\
%& \\
\bottomrule
\end{tabular}
\end{table*}
\endgroup