### 2011-09-23 expectationPeopleTable.tex
table with data

\begingroup
\newcommand{\add}{\hspace{0pt}}

\definecolor{white}{rgb}{1,1,1}
\definecolor{mygray}{rgb}{0.7,0.7,0.7}
\definecolor{yourgray}{rgb}{0.4,0.4,0.4}
\definecolor{black}{rgb}{0.0,0.0,0.0}

\newdimen\qdx
\newdimen\qda
\newdimen\qdb
\newdimen\qd
\def\rrrr#1#2#3#4#5#6{\qd=#4 % length of bar for 1.0
\qdx=\qd\multiply\qdx by 5\divide\qdx by 4
\qda=\qd
\qdb=\qd
\multiply\qda by #1\divide\qda by #3\multiply\qdb by #2\divide\qdb by #3\advance\qdb by -\qda
    \leavevmode\hbox to \qdx{\hfil\vbox{%
    \hbox{\vrule\vbox{\hrule\hbox to 1\qd
            {\vrule depth0pt height#6 width \qda#5\vrule depth0pt height#6 width \qdb\hfill}\hrule}\vrule}
    }\hfil}}
\def\rrr#1#2#3#4{\rrrr{#1}{#2}{#3}{0.15cm}{#4}{1.5ex}}

\def\w{\rrr{0}{1}{1}{\color{white}}}
\def\l{\rrr{0}{1}{1}{\color{mygray}}}
\def\g{\rrr{0}{1}{1}{\color{yourgray}}}
\def\b{\rrr{0}{1}{1}{\color{black}}}
\def\0{\w}
\def\1{\l}
\def\2{\b}
\def\angle{90}

\begin{table*}[tb]
\centering
%\footnotesize
%\scriptsize
\begin{tabular}{l@{\hspace{-3cm}}rc@{\hspace{5pt}}c@{\hspace{5pt}}c@{\hspace{5pt}}c@{\hspace{5pt}}c@{\hspace{5pt}}c@{\hspace{5pt}}c@{\hspace{5pt}}c@{\hspace{5pt}}c@{\hspace{5pt}}c}
\toprule
\vspace{1.35cm}\\
& Expectations 
&\multicolumn{1}{c}{\begin{rotate}{\angle}SW Directors\end{rotate} }
&\multicolumn{1}{c}{\begin{rotate}{\angle}Tech Leads\end{rotate} }
&\multicolumn{1}{c}{\begin{rotate}{\angle}Team Leads\end{rotate} }
&\multicolumn{1}{c}{\begin{rotate}{\angle}Developers\end{rotate} }
&\multicolumn{1}{c}{\begin{rotate}{\angle}Qa \& Ops\end{rotate} }
&\multicolumn{1}{c}{\begin{rotate}{\angle}Prod Mgmt\end{rotate} }
&\multicolumn{1}{c}{\begin{rotate}{\angle}Project Mgmt\end{rotate}} 
&\multicolumn{1}{c}{\begin{rotate}{\angle}Support \& Field\end{rotate}} \\
%& Expectations & \multicolumn{2}{l}{People} \\
\midrule
%&&& $p_1$ & $p_2$& $p_3$& $p_4$& $p_5$& $p_6$& $p_7$& $p_8$& $p_9$& $p_{10}$& $p_{11}$& $p_{12}$& $p_{13}$& $p_{14}$& $p_{15}$& $p_{16}$& $p_{17}$& $p_{18}$& $p_{19}$& $p_{20}$& $p_{21}$& $p_{22}$& $p_{23}$& $p_{24}$& $p_{25}$& $p_{26}$& $p_{27}$& $p_{28}$& $p_{29}$& $p_{30}$& $p_{31}$& $p_{32}$& $p_{33}$& $p_{34}$& $p_{35}$& $p_{36}$\\
%&&&p&p&p&p&p&p&p&p&p&p&p&p&p&p&p&p&p&p&p&p&p&p&p&p&p&p&p&p&p&p&p&p&p&p&p&p\\
%\midrule
\textbf{Technical - Core} 
&Write bug-fix code& \0\0&\1\1\1\1&\1\2\0\2\2&\1\1\1\0\1\0\2\1&\0\0\0\0&\0\0\0\0&\0\0\0&\0\0\0\0\0\0 \\
&Maintain legacy code& \0\0&\0\1\2\1&\2\0\0\2\2&\2\0\2\0\2\2\0\0&\0\0\0\0&\0\0\0\0&\0\0\0&\0\0\0\0\0\0  \\
&Write new-feature code& \0\0&\2\2\1\1&\1\2\0\2\1&\1\2\0\0\0\1\2\2&\0\0\0\0&\0\0\0\0&\0\0\0&\0\0\0\0\0\0 \\
&Write customization (or script) code& \0\0&\0\0\0\0&\0\0\2\0\0&\0\0\0\2\0\0\0\0&\0\0\0\0&\0\0\0\1&\0\0\0&\0\2\0\0\0\0\\
%& \\
%
%
\midrule
\textbf{Technical - Periphery}
& Design user interface&\0\0&\2\1\0\0&\0\0\0\0\0&\0\0\0\0\0\0\0\2&\0\0\0\0&\1\2\0\0&\0\0\0&\0\0\0\0\0\0\\
& Write documentation&\0\0&\0\1\0\1&\0\0\1\0\1&\1\0\1\1\0\0\0\0&\1\0\1\0&\2\0\2\0&\0\0\0&\0\0\0\0\0\0\\
&Write low-level (or unit) test code&\0\0&\1\1\0\1&\1\1\1\1\1&\1\1\0\1\1\1\0\0&\0\0\0\0&\0\0\0\0&\0\0\0&\0\0\0\0\0\0\\
& Write Build Code and prepare a release&\0\0&\0\0\0\0&\0\0\0\1\0&\0\0\0\0\0\0\0\0&\0\0\2\0&\0\0\0\0&\0\0\0&\0\0\0\0\0\0\\
& Write experimental (or exploratory) code&\0\0&\1\2\0\2&\0\0\0\0\0&\0\0\0\0\2\0\1\0&\0\0\0\0&\0\0\0\0&\0\0\0&\0\0\0\0\0\0\\
& Write high-level (or integration) test code&\0\0&\0\0\1\0&\0\0\0\2\0&\0\0\0\0\0\0\0\0&\2\1\0\0&\0\0\0\0&\0\0\0&\0\0\0\0\0\0\\
& Write configuration and implementation code&\0\0&\0\0\0\0&\0\0\1\0\0&\0\0\0\0\0\0\0\0&\0\0\0\0&\0\0\0\0&\0\0\0&\0\2\0\0\0\0\\
& Write and maintain tools to aid in development&\0\0&\0\0\0\1&\0\0\0\1\0&\0\0\0\0\0\0\0\0&\0\0\2\1&\0\0\0\0&\0\0\0&\0\0\0\0\0\0\\
%& \\
%
%
\midrule
\textbf{Technical - High level}
& Design system architecture&\0\0&\2\2\2\2&\0\0\0\1\1&\0\0\0\0\0\0\0\1&\0\0\0\0&\1\0\0\0&\0\0\0&\0\0\0\0\0\0\\
& Elicit and document requirements&\0\0&\0\0\0\1&\0\0\0\1\1&\0\0\0\0\0\0\0\0&\0\0\0\0&\2\2\2\0&\0\0\0&\0\0\0\2\1\2\\
& Determine if product meets quality goals&\1\0&\1\1\0\0&\1\0\0\1\0&\0\1\0\0\0\0\1\1&\2\2\1\2&\0\2\0\1&\0\0\0&\0\0\0\0\1\1\\
& Set roadmap for system (prioritize requirements)&\1\0&\1\0\0\1&\1\0\0\2\0&\0\0\0\0\0\0\0\0&\0\0\0\0&\0\2\2\2&\0\0\0&\0\0\0\0\2\1\\
& Provide domain context to non-domain experts (user proxy)&\0\0&\0\0\0\0&\0\0\0\0\0&\0\0\0\0\0\0\0\0&\0\0\0\0&\0\2\0\2&\0\0\0&\0\2\0\2\2\0\\
%& \\
%
%
\midrule
\textbf{Technical - Expertise}
& Tackle the toughest technical issues and bugs&\0\0&\0\2\1\2&\2\0\2\0\0&\1\0\0\0\2\0\0\0&\0\0\0\0&\0\0\0\0&\0\0\0&\0\0\0\0\0\0\\
& Provide historical perspective on design decisions&\0\0&\2\1\0\2&\0\0\0\0\0&\2\0\2\0\0\0\0\0&\0\1\0\0&\2\0\0\0&\0\0\0&\0\0\0\0\0\0\\
& Provide guidance (mentorship) to newcomers on technical issues&\0\0&\1\0\0\0&\0\0\2\2\2&\2\0\0\0\0\0\0\1&\0\0\0\0&\0\0\0\0&\0\0\0&\0\0\0\0\0\0\\
%& \\
%
%
\midrule
\textbf{Technical - Management}
& Determine lifecycle process and practices&\2\1&\1\0\0\0&\1\0\1\2\1&\0\0\0\0\0\0\0\0&\0\0\0\0&\0\0\0\0&\1\2\0&\0\0\0\0\0\0\\
& Assign low-level tasks (tickets) to individuals&\0\1&\1\0\0\0&\2\0\1\1\0&\0\0\0\0\0\0\0\0&\0\0\0\2&\0\0\1\0&\1\1\0&\1\2\1\0\0\0\\
& Assign high-level tasks (projects) to individuals&\2\2&\0\0\0\1&\0\0\0\0\0&\0\0\0\0\0\0\00&\0\0\0\0&\0\0\1\0&\0\0\0&\0\0\0\0\0\0\\
& Coordinate a product release or implementation&\1\1&\1\0\1\1&\2\0\0\2\0&\0\0\0\0\0\0\0\0&\0\1\2\2&\1\2\2\0&\2\2\1&\0\1\0\0\0\0\\
& Provide estimates on remaining or potential tasks&\1\1&\2\1\1\0&\2\0\1\2\0&\0\0\0\0\0\0\0\0&\0\0\0\0&\0\1\1\0&\2\2\1&\0\0\0\0\0\0\\
& Connect an issue with the person with appropriate expertise&\1\1&\0\0\0\1&\2\0\0\0\0&\0\0\0\0\0\0\0\0&\0\0\0\0&\0\0\0\0&\1\1\0&\1\2\2\0\1\0\\
& Handle obstacles and interruptions for technical teams (PoC)&\1\2&\0\0\0\0&\2\1\1\0\1&\0\0\0\0\0\0\0\0&\0\0\0\2&\0\0\0\0&\2\2\2&\0\0\0\0\0\0\\
%& \\
%
%
\midrule
\textbf{Organizational - External Interface}
& Receive and respond to user feedback&\0\2&\0\0\0\0&\0\1\0\1\1&\0\0\0\0\0\0\0\1&\0\0\0\1&\0\0\0\2&\2\2\0&\2\2\2\0\1\1\\
& Obtain new users or sources of funding&\0\2&\0\0\0\0&\0\0\0\0\0&\0\0\0\0\0\0\0\0&\0\0\0\0&\0\0\0\1&\0\0\0&\0\0\0\2\0\0\\
& Provide domain legitimacy to organization&\0\0&\0\0\0\0&\0\0\0\0\0&\0\0\0\0\0\0\0\0&\0\0\0\0&\0\2\0\2&\0\0\0&\0\0\0\0\0\0\\
& Provide technical legitimacy to organization&\2\1&\2\0\0\2&\0\0\0\0\1&\0\0\0\0\0\0\0\0&\0\0\0\1&\0\0\0\0&\0\0\0&\0\0\0\0\0\0\\
& Advocate for the resolution of user/customer problems&\1\2&\0\0\0\0&\1\0\0\0\0&\0\0\0\0\0\0\0\0&\0\0\0\1&\0\1\0\0&\2\2\0&\2\2\2\1\2\2\\
& Represent the organization's interests to funders/customers&\0\2&\0\0\0\1&\0\1\0\0\0&\0\0\0\0\0\0\0\0&\0\0\0\0&\0\1\0\2&\0\0\0&\0\0\0\1\1\1\\
%& \\
%
%
\midrule
\textbf{Organizational - Management}
& Deal with personnel issues&\2\2&\0\1\0\0&\1\1\0\0\0&\0\0\0\0\0\0\0\0&\0\0\0\1&\0\0\0\1&\0\0\2&\0\0\1\0\0\0\\
& Deal with otherwise unattended issues&\1\1&\0\0\0\1&\1\0\0\0\0&\0\0\0\0\0\0\0\0&\0\2\0\1&\0\1\0\1&\0\0\2&\0\0\0\0\0\0\\
& Set up and modify organizational structure&\2\2&\0\0\0\1&\0\0\0\0\0&\0\0\0\0\0\0\0\0&\0\0\0\1&\0\0\1\1&\0\0\0&\0\0\1\0\0\0\\
& Communicate organizational status and news&\2\2&\0\0\0\0&\1\0\1\0\0&\0\0\0\0\0\0\0\0&\0\0\0\1&\0\0\0\0&\1\1\1&\0\0\1\0\0\0\\
& Deal with financial, clerical, or system administration issues&\0\1&\0\0\0\0&\0\0\0\0\0&\0\0\0\0\0\0\0\0&\0\0\0\1&\0\0\0\0&\0\0\2&\0\0\0\0\0\0\\
%& \\
%
%
\midrule
\textbf{Organizational - Strategic}
& Determine long-term vision for products&\1\2&\1\0\0\2&\0\0\0\0\0&\0\0\0\0\0\0\0\0&\0\0\0\0&\0\2\0\2&\0\0\0&\0\0\0\0\0\0\\
& Determine emphasis on new market areas&\0\1&\0\0\0\1&\0\0\0\0\0&\0\0\0\0\0\0\0\0&\0\0\0\0&\0\2\0\1&\0\0\0&\0\0\0\0\0\0\\
%& \\
\bottomrule
\end{tabular}
\end{table*}
\endgroup